\section{Conclusion}

Knowledge arbitration should be treated as an inference problem, not as a
retrieval heuristic. Once written as a posterior-predictive decision problem,
the right comparison is not between one tuned adaptive rule and another, but
between reliability-aware arbitration and fixed trust policies that cannot
adapt to conflict families. Our corrected benchmark waves show that the
Bayes-style rule wins that comparison cleanly, while fixed trust in retrieval
or fixed trust in parametric memory incurs enormous regret.

The reasoning-time result is also now clearer than it was at the start of the
project. The original monotone claim does not survive contact with real traces.
What survives is a more interesting law: reasoning amplifies overconfidence
under conflict, and the shape of that amplification depends on the conflict
family. WikiContradict exhibits an intermediate-CoT peak with partial recovery.
ConflictBank conflict becomes even harsher at 14B, remaining severely
overconfident through long CoT. The completed same-family Qwen sweep shows that
this should not be oversold as a universal 7B-vs-14B asymmetry: Qwen does not
recover on ConflictBank through 32B, while 32B is the first scale at which
recovery reappears on WikiContradict. The finished theorem-3 claim is therefore
benchmark-dependent two-regime behavior, not a family-invariant scale law. The
cleanest interpretation is an RLVR-conditioned misspecification story:
reasoning-time confidence can improve in low-conflict regimes and collapse in
controlled conflict regimes where endogenous evidence keeps sharpening the wrong
answer.

The immediate next steps are now straightforward. The theory side needs full
proof writeups for the Bayes-optimal rule and the fixed-policy minimax lower
bound, while the theorem-3 side should continue to be written as a structural
misspecification account rather than as a theorem of equal formal status. The
empirical side still needs direct training-time comparisons to RLCR-style
reward redesign, broader checkpoint-family sweeps, and mitigation experiments
such as bagged reasoning and confidence decoupling. But the main paper arc is
no longer speculative: fixed trust policies are wrong, a reliability-aware
rule does better, and real reasoning traces reveal a structured calibration
pathology under knowledge conflict. The associated method story is now concrete
too: confidence-only $\eta$-tempering is enough to nearly recalibrate
naturalistic contradiction, but not enough to rescue answer-level
controlled-conflict failures.

\paragraph{Scope and limitations.}
Three practical boundaries matter for interpretation. First, the current paper
does not yet report a completed head-to-head number against RLCR
\citep{damani2025rlcr}; our training-time evidence is still lighter-weight than
a full reward-level redesign on the same backbone. Second, the retrieval demo
is deployment-motivated but not production-scale: it uses bounded
Wikipedia-backed retrieval rather than a continuously refreshed live stack.
Third, the benchmark, prompt, and calibration suite here is English-only.
