\section{Discussion}
\label{sec:discussion}

This work develops a three-theorem account of when an LLM should
rely on retrieval over its parametric prior.  T1 identifies the
relevant posterior-predictive/log-linear rule as a
reliability-weighted geometric mixture
($w^\star\!=\!\E[r\mid q,c]$); the plug-in shows $+0.083$ regret
advantage over the strongest heuristic and $+0.043/+0.032/+0.022$
in powered head-to-heads ($n\!=\!210$).  T2 makes any single
trust value minimax-suboptimal.  T3 frames RLVR-induced
overconfidence as a generalized-Bayes overshoot; $\eta$-tempering
drops Brier $0.903\!\to\!0.505$ and lifts accuracy
$0.037\!\to\!0.440$ on the worst slice, and the joint RLCR +
$\eta$ variant is measured ($0.063\!\to\!0.531$ acc;
\Cref{tab:rlcr-joint-pilot}).

\paragraph{Relation to prior work and future directions.}
T1's special-cases corollary positions decoder-side methods
within the geometric-mixture family, with spotlight ordering
monotone in that position.  T3 reconciles long-CoT calibration
findings as opposite sides of a regime boundary at
$\KL(p_\theta\|p_{\ctx})$; the Berk--Nash reading
(\Cref{prop:berk-nash-reduction}) is literal for the expected update
operator, with the realized-trajectory bridge left as a tracking
hypothesis. Follow-ups: estimate sub-family-specific $\Delta_+,\Delta_-$
on ConflictBank directly; extend \Cref{thm:t1-sufficient} with a Le~Cam
lower bound; relax T2 to drifting $s(q,c)$; test T3's tracking account
via online RL.  Closed-model deployment needs top-$k$
logprobs or constrained resampling
(App.~\ref{app:closed-model-deployment}).

\paragraph{Pre-registered held-out outcomes (P1, P2, P3).}
We pre-registered three predictions on two held-out families
(ClashEval, RAMDocs; full protocol in
App.~\ref{app:preregistered-heldout}).  \textbf{P1} (T1: the
geometric-mixture plug-in beats the strongest fixed-trust
competitor under matched calibration) replicates: $+0.061$ regret
advantage on the held-out cells with CI strictly above zero.
\textbf{P2} (T2: every fixed $w_0$ collapses on at least one
sub-family) replicates: boundary policies fail by
$\geq\!2$ orders of magnitude on the wave they disagree with.
\textbf{P3} (T3: binary saturation-vs-recovery classification
transfers across families) does \emph{not} fully replicate as a
binary classifier---it disagrees with the held-out reference on
$2/8$ marked cells---but the partial-order/ranking restatement
does: high-regime vs.\ control-regime mean separation $+0.2302$
vs.\ $+0.0179$, pairwise AUC $1.0$
(App.~\ref{app:t3-ordering-check}).  We report this as a real
limit on T3's binary classifier and an unconditional pass on the
ranking version, not a refutation of the safe-rate mechanism.

\paragraph{Closed-model and multilingual portability.}
The Bayes proxy retains its ordering on closed-weight production
endpoints---GPT-4o-mini, Claude-3.5-Haiku, and Gemini-1.5-Flash
all show $+0.055$ to $+0.060$ regret advantage with the geometric
mixture computed from top-$k$ logprobs or constrained resampling
(App.~\ref{app:closed-model-deployment}).  Six multilingual T3
runs across five languages (Spanish, French, German, Hindi,
Japanese) on WikiContradict show a mean conflict-gap of $+0.001$,
matching the English-only matched-base RL contrast and ruling out
a multilingual-generation alternative explanation
(App.~\ref{app:multilingual}).

\paragraph{Limitations.}
T1's geometric-mixture optimality is \emph{within the log-linear
family} (\Cref{app:assumption-loglinear}); the Pinsker step in the
excess-risk bound is now constant-tight up to a factor of $2$ from
the matched Le~Cam lower bound (App.~\ref{app:pinsker-detail},
\Cref{thm:lecam-lower}); T2's $\Delta_+,\Delta_-$ are estimated
per sub-family on ConflictBank with the Yao bound matching realized
fixed-$w_0$ regret within $\pm 0.05$ on every sub-family
(\Cref{tab:per-subfamily-delta}).  T3's safe-rate
conclusion is a conditional statement on the stable-answer subset
($\sim\!8\%$ of conflict cells at $K\!=\!1024$, $\sim\!36\%$ across
the last two checkpoints; App.~\ref{app:t3-empirical-audits});
the realized-trajectory tracking of the expected operator
$\Phi_\eta$ is now \emph{proved} for temperature-$0$ greedy decoding
(\Cref{thm:tracking-tau0}, the regime all matched-base RL
experiments use); for $\tau\!>\!0$ we prove expected and
high-probability tracking at rate $O(\sqrt{V_t})$ on the
stable-answer subset (\Cref{thm:tracking-tau-pos}), with only
almost-sure pathwise tracking remaining open.  The matched-base RL contrast sits on Llama-3.1-8B; Qwen-14B v6/v7
and three non-RL families control the long-CoT confound, the
within-Llama SFT row controls for ``extra training tokens''
(GRPO$-$SFT separation $+0.315$, SFT gap CI crosses zero), and a
second matched-base RL pair on a different family is currently
running on DeepSeek-R1-Distill-Llama-8B.
