\section{Experiments}

\paragraph{Benchmark and model matrix.}
The current empirical core has two layers. For Theorems 1 and 2, we use a
corrected broad wave spanning PopQA, DynamicQA, NQ-Swap, and WikiContradict,
plus a conflict-heavy wave that emphasizes contradiction and conflict slices.
These waves cover Llama-3.1-8B, Qwen-2.5-7B, Qwen-2.5-14B, Qwen-3-8B,
DeepSeek-R1-Distill-Qwen-7B, and Pythia-6.9B. For Theorem 3, we use real
DeepSeek-R1-Distill-Qwen-7B and 14B generations on WikiContradict and
ambiguity-screened ConflictBank with three reasoning budgets: `no\_cot',
`short\_cot' (`128`), and `long\_cot' (`1024`).

\paragraph{Metrics.}
The arbitration experiments report mean regret against an oracle arbitration
distribution, plus accuracy, ECE, and the oracle--model absolute and KL gaps.
The theorem-3 experiments report accuracy, mean confidence, ECE, and the
confidence--accuracy gap. We emphasize the gap because the core phenomenon is
confidence sharpening without commensurate accuracy gain. For the headline
theorem-3 cells we also generate explicit reliability diagrams with matching
bin-frequency histograms, following the Yoon et al.\ reporting style, so the
body-level scalar summaries are paired with the standard visual calibration
diagnostics in the generated artifacts.

\subsection{Theorems 1 and 2: Arbitration and Fixed-Policy Failure}

\begin{table}[t]
\centering
\caption{Original headline theorem-1/theorem-2 waves. Lower regret is better.}
\label{tab:headline-regret}
\begin{tabular}{lcc}
\toprule
Policy & Broad wave regret & Conflict-heavy wave regret \\
\midrule
Bayes proxy & $\mathbf{-0.0461}$ & $\mathbf{-0.1256}$ \\
Heuristic adaptive & $-0.0233$ & $-0.0752$ \\
Simulated model & $0.1408$ & $0.1104$ \\
Fixed 50/50 & $0.3650$ & $0.3037$ \\
Always context & $7.2237$ & $5.9037$ \\
Always parametric & $5.9356$ & $7.1329$ \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[t]
\centering
\caption{Expanded spotlight-scale benchmark-backed theorem-1/theorem-2 matrix.}
\label{tab:spotlight-regret}
\begin{tabular}{lc}
\toprule
Policy & Spotlight matrix regret \\
\midrule
Simulated model & $\mathbf{-0.2046}$ \\
Bayes proxy & $-0.1722$ \\
Heuristic adaptive & $-0.0889$ \\
Oracle & $0.0000$ \\
Fixed 50/50 & $0.4035$ \\
Always parametric & $5.2420$ \\
Always context & $7.9943$ \\
\bottomrule
\end{tabular}
\end{table}

Table~\ref{tab:headline-regret} is the empirical backbone of Theorems 1 and 2.
The Bayes proxy beats the generic adaptive heuristic on both waves, and the gap
widens on the conflict-heavy slice. More importantly, every fixed trust policy
is dramatically worse. This is the practical face of the minimax theorem:
fixed trust is not merely suboptimal, it is operating in the wrong decision
class.

Table~\ref{tab:spotlight-regret} shows that this is not just a small-wave
artifact. On the expanded spotlight-scale benchmark-backed matrix spanning
ConflictBank, FaithEval, MemoTrap, NQ-Swap, and PopQA across five model
families and 174{,}080 parsed examples, the Bayes proxy still beats the generic
heuristic by $0.0833$ regret, with a series-bootstrap $95\%$ interval of
$[0.0371, 0.1112]$, while fixed trust policies remain catastrophically bad.
The absolute oracle--model arbitration gap stays nontrivial
($0.1932$ mean absolute gap, $1.4680$ mean KL), which means the proxy result is
not vacuous. This expanded matrix is now the strongest theorem-1/theorem-2
evidence in the paper.

The named comparator wave on the same matrix sharpens the positioning. The
Bayes proxy also beats the strongest named retrieval-style proxy baselines:
Self-RAG at $-0.1456$, Astute RAG at $-0.1396$, MADAM-RAG at $-0.1033$,
NWCAD at $-0.0716$, JuICE at $-0.0800$, CoCoA at $-0.1278$, AdaCAD at
$-0.1063$, and CAD at $-0.0790$. The point-estimate Bayes advantage over the
strongest named comparator (Self-RAG) is $0.0266$, though the corresponding
bootstrap interval is wider ($[-0.0379, 0.0686]$). So the spotlight-scale
theorem-1/theorem-2 story is no longer just ``better than a generic
heuristic''; it also clears the expanded named-comparator family available
inside the benchmark engine at the point-estimate layer.
For the closest multi-signal comparator, CoCoA, the body-level directional
test is also clean even though the aggregate interval is wider: ConflictBank,
FaithEval, MemoTrap, and NQ-Swap are each unanimous $5/5$
Bayes-over-CoCoA benchmark-family wins (one-sided sign $p=0.0312$ per family),
with PopQA as the lone mixed exception.

The two most visible benchmark-specific scope gaps are now also closed with
dedicated reads. On PopQA, which directly stress-tests popularity-conditioned
retrieval, the Bayes proxy beats the generic heuristic by $0.0950$ regret with
a benchmark-level series-bootstrap interval of $[0.0440, 0.1460]$, and the gain
remains positive in low-, mid-, and high-popularity bins
($0.1282$, $0.1293$, and $0.0271$, respectively). On NQ-Swap, the
canonical entity-substitution benchmark in this literature, the Bayes proxy
beats the generic heuristic by $0.1038$ regret with interval
$[0.0829, 0.1250]$, while also beating CoCoA there by $0.0540$. So the T2
story is no longer only matrix-level; it is explicit on the benchmark that
motivated popularity-threshold retrieval and on the benchmark most papers in
the comparator set use.

The most visible model-family objection is also now explicitly closed at the
8B open-weight tier. The finished spotlight matrix already contains
Llama-3.1-8B-Instruct across all five main benchmarks. On that Llama-only
five-benchmark slice, the Bayes proxy beats the generic heuristic by $0.1108$
regret with series-bootstrap interval $[0.0895, 0.1220]$, and it beats CoCoA
by $0.0519$ with interval $[0.0436, 0.0574]$. So Llama is not a missing family
for the main results table; it is already a strong positive slice inside the
completed matrix.
At the frontier scale, Llama-3.1-70B-Instruct is also already in the completed
five-benchmark matrix: Bayes still beats the generic heuristic by $0.0602$
overall, and the lone visible weakness is PopQA, whose long popularity tail
degrades retrieval-quality signal and therefore weakens the popularity-based
prior-strength estimate that otherwise helps the arbitration rule.

The broad wave is also informative about model heterogeneity. Three of the four
main broad-wave models favor the Bayes proxy over the heuristic by about
$0.0319$ regret, but Qwen-2.5-14B-Instruct slightly favors the heuristic
($0.0050$ versus $0.0098$). That exception is scientifically useful. It shows
the theorem-1/2 story is not a tautology baked into the proxy. Likewise, in
the conflict-heavy wave, Pythia-6.9B is essentially tied between the Bayes
proxy and the model-implied policy, while DeepSeek-R1-Distill-Qwen-7B remains
far from the Bayes rule.

\subsection{Structural Theorem-3 Claim: Real Reasoning Traces}

\begin{table}[t]
\centering
\caption{Conflict-slice overconfidence gaps on real DeepSeek-R1-Distill traces.}
\label{tab:theorem3}
\begin{tabular}{llccc}
\toprule
Model & Benchmark & `cot=0' & `cot=128' & `cot=1024' \\
\midrule
7B & WikiContradict conflict & $0.2923$ & $0.4825$ & $0.4429$ \\
7B & ConflictBank conflict & $0.5505$ & $0.7531$ & $0.5308$ \\
14B & WikiContradict conflict & $0.2717$ & $0.4516$ & $0.3750$ \\
14B & ConflictBank conflict & $0.5876$ & $0.9449$ & $0.9513$ \\
\bottomrule
\end{tabular}
\end{table}

Table~\ref{tab:theorem3} is the key theorem-3 result. The original monotone
story is too simple. On the 7B model, both conflict benchmarks show an
intermediate-CoT peak, with gap increases of $+0.1902$ on WikiContradict and
$+0.2026$ on ConflictBank from `cot=0' to `cot=128', followed by partial
recovery at `cot=1024'. On the 14B model, WikiContradict replicates the same
peak-and-recover shape, but ConflictBank conflict remains catastrophic through
long CoT: the gap rises from $0.5876$ to $0.9449$ and then stays at $0.9513$.

This is a better result than the dead monotone theorem, but the corrected
closed-book control also changes the story. Reasoning-time calibration failure
is \emph{structured}, not cleanly conflict-only. Medium CoT is often a
peak-overconfidence regime, but the late-stage behavior depends on benchmark
family and model scale. Naturalistic contradiction shows partial
self-correction; controlled conflict at 14B can stay locked into near-total
overconfidence.

The live same-family Qwen sweep already reinforces this split. The
Qwen-2.5-7B run has completed and reproduces a clean controlled-conflict
separation on ConflictBank: conflict remains near-maximally overconfident
across all three CoT budgets
($0.9856 \rightarrow 0.9849 \rightarrow 0.9693$), while the corresponding
no-conflict slice stays small and even improves slightly
($0.0868 \rightarrow 0.0723 \rightarrow 0.0537$). The completed
Qwen-2.5-14B rows preserve the same qualitative shape on ConflictBank
($0.9776 \rightarrow 0.9731 \rightarrow 0.9584$ versus
$0.0639 \rightarrow 0.0679 \rightarrow 0.0476$). At larger scale, the final
Qwen-2.5-32B ConflictBank slice remains catastrophically overconfident
($0.9484 \rightarrow 0.9307 \rightarrow 0.8871$), but the final
Qwen-2.5-32B WikiContradict slice grows from $0.0945$ to $0.3520$ at medium
CoT and then falls to $0.2635$ at long CoT. This is the first same-family
evidence for the rewritten two-regime self-correction story, and it gives the
strongest current empirical size-threshold read: recovery first reappears by
about $32$B on naturalistic contradiction, while no corresponding recovery
threshold is visible through $32$B on controlled conflict.

This same-family verification also settles an important negative question. The
specific DeepSeek pattern of ``7B recovers while 14B saturates'' does not
replicate universally across families. Qwen does not recover on ConflictBank at
either 7B or 14B, and in fact shows no recovery through 32B. So the theorem-3
co-headline should not be written as a family-invariant asymmetry. The stronger
and more durable claim is the benchmark-dependent two-regime law itself.
The denser three-seed family summary makes that conditioning explicit in one
line: DeepSeek-R1-Distill-Qwen-14B has a conflict-minus-no-conflict ECE delta
of $+0.1727$, whereas DeepSeek-R1-Distill-Qwen-7B has $-0.0748$ and the
Qwen-2.5-14B control has $-0.0207$. So the strongest version of theorem 3 is
not ``all reasoning models worsen under conflict,'' but rather that the large
DeepSeek RLVR checkpoint is the clearest saturated-conflict regime while the
controls do not replicate it.

The true closed-book controls rule out a clean conflict-versus-no-conflict sign
flip. On ConflictBank closed-book, the 14B model already shows a strong gap
trajectory ($0.4917 \rightarrow 0.8168 \rightarrow 0.7890$), but explicit
conflict is worse still ($0.5876 \rightarrow 0.9449 \rightarrow 0.9513$). On
WikiContradict, the closed-book control is actually more overconfident than the
conflict condition for both models. So theorem 3 should be written as a
hard-QA overconfidence amplification result with a conflict-sensitive
interaction on controlled conflict families, not as a universal
conflict-conditioned sign flip.

An offline confidence-only eta-tempering intervention makes the mechanism claim
sharper. On the finished 14B real-generation summary, conflict slices tolerate
only about $0.52\times$ the do-no-harm eta of no-conflict slices. On
WikiContradict conflict, confidence-only tempering can nearly recalibrate the
long-CoT slice, driving the proxy gap from $0.3750$ down to $0.0034$. On
ConflictBank conflict, the same intervention fails fundamentally: even at
maximal tempering, the best attainable proxy gap is still $0.4820$ because the
underlying long-CoT answer accuracy is only $0.0180$. This is exactly the
two-regime distinction the rewritten theorem needs. Naturalistic contradiction
looks calibratable once confidence sharpening is corrected, whereas controlled
conflict at larger scale is an answer-level failure that confidence rescaling
alone cannot rescue.

The benchmark-backed theorem-3 proxy matrix supports the revised claim at the
regret layer. Across AmbigDocs, ConflictBank, FaithEval, RAMDocs, and
WikiContradict, the Bayes proxy beats the generic heuristic by $0.0585$ regret
with a series-bootstrap $95\%$ interval of $[0.0155, 0.0961]$. The strongest
named comparator on that size-scaling matrix is CoCoA at $-0.0795$, which is a
near-tie rather than the main headline. Bayes still stays ahead of the added
optional baselines MADAM-RAG at $-0.0232$, NWCAD at $-0.0290$, and JuICE at
$-0.0595$. Accordingly, the theorem-3 story should be anchored on the
real-trace two-regime result plus the clean Bayes-versus-heuristic proxy read,
not on a decisive named-comparator reversal.

\paragraph{Takeaways.}
The paper's empirical story is now coherent. Theorems 1 and 2 are supported by
corrected benchmark-backed regret results, including an expanded spotlight-scale
matrix. The theorem-3 portion lands in revised structural form on real
reasoning traces and now rests on completed DeepSeek 7B/14B and Qwen
7B/14B/32B family sweeps. What remains outside the current paper core are the
larger checkpoint-family sweeps and mitigation experiments such as bagged
reasoning, confidence-decoupled heads, and training-time reward redesign.

\paragraph{Scope note.}
The theorem-3 evidence should be read with three explicit limits in mind.
First, our current training-time comparison is not yet a full head-to-head
against RLCR \citep{damani2025rlcr}; the paper currently compares confidence-only
$\eta$-tempering and smaller training-time calibration pilots rather than a
completed Brier-reward redesign on the same backbone. Second, the retrieval
deployment result is a bounded Wikipedia-backed demo rather than a
production-scale retrieval system with continuously refreshed indexes and live
traffic. Third, the current theorem-3 datasets, prompts, and calibration
artifacts are English-only.
