\section{Related Work}
\label{sec:related}

\paragraph{Knowledge conflict in LLMs.}
Surveys carve up knowledge conflict into context-memory, inter-context, and
intra-memory sub-types~\citep{xu2024knowledgeconflicts,xie2024conflictqa}.
Benchmarks span synthetic counter-memory across $7$k entities and three
conflict types~\citep{su2024conflictbank}, real Wikipedia
contradictions~\citep{hou2024wikicontradict}, temporal
drift~\citep{marjanovic2024dynamicqa}, popularity
stratification~\citep{mallen2023popqa}, and entity
substitution~\citep{xie2024conflictqa}. These efforts diagnose the problem;
we ask what the optimal arbitration rule looks like and identify
(\Cref{thm:bayes-arbitration}) the relevant posterior-predictive object
inside the log-linear family used by decoder-side methods.

\paragraph{Adaptive RAG and decoder-level arbitration.}
PopQA-style retrieval thresholds on
popularity~\citep{mallen2023popqa}; Self-RAG learns when to
retrieve~\citep{asai2024selfrag}; FLARE retrieves on
demand~\citep{jiang2023flare}; CRAG corrects retrieval
errors~\citep{yan2024crag}; Astute~RAG and MADAM-RAG
debate conflicting documents~\citep{wang2024astuterag,wang2024madamrag}.
At the decoder level, CAD~\citep{shi2024cad}, AdaCAD~\citep{wang2025adacad},
CoCoA~\citep{khandelwal2025cocoa}, NWCAD~\citep{khandelwal2024nwcad}, and
JuICE~\citep{cohen2024juice} reweight contextual logits with hand-crafted
signals. \Cref{cor:special-cases} shows where each sits relative to the
Bayes optimum: CAD is $w\!=\!2$ (over-amplification); AdaCAD's JSD weight
is the first-order Laplace approximation of $\E[r\mid c]$ under a Beta
prior with Hellinger-affinity likelihood
(\Cref{lem:adacad-laplace} in App.~\ref{app:thm1}); CoCoA is a
non-Bayesian ensemble of three reliability proxies that need not converge
to $w^\star$ without additional calibration assumptions.

\paragraph{Reasoning-time calibration.}
The recent literature is mutually contradictory at first glance.
\citet{yoon2025confidence} report that reasoning models express confidence
\emph{better} than chat models on standard QA;
\citet{halawi2024overthinking} document depth-wise overthinking;
\citet{lacombe2025dontthinktwice} find more reasoning hurts calibration on
knowledge-intensive tasks; \citet{welch2026cost} isolate ``implicit answer
conditioning'' in VLMs; \citet{damani2025rlcr} show that binary-reward RL
(GRPO~\citep{shao2024grpo}, of which DeepSeek-R1~\citep{deepseek2025r1} is
the canonical instance) degrades calibration and that Brier-style rewards
repair it; \citet{bereket2025grpo} document GRPO-induced overconfidence;
\citet{sun2025context} show rationales shift weight onto context.
\Cref{thm:eta-bayes} reconciles this literature: in the well-specified or
low-conflict regime ($\eta\!<\!\bar\eta$) reasoning helps; when conflict
reduces the safe rate enough, the same update can cross into an
overshoot-driven saturation regime.

\paragraph{Misspecified Bayes as a template.}
The mechanism rests on classical results: Berk's KL-projection
concentration~\citep{berk1966}; Kleijn--van der Vaart's nonparametric
extension~\citep{kleijn2006}; Bissiri--Holmes--Walker's coherent
generalized-Bayes updates with learning rate
$\eta$~\citep{bissiri2016}; Gr\"unwald--van Ommen's safe rate
$\bar\eta$~\citep{grunwald2017}; and the endogenous-evidence Berk--Nash
results of \citet{esponda2016,fudenberg2021}. We do not reprove these or
claim the Berk--Nash machinery as novel; we transport it to the LLM CoT
setting where the trace is generated by the very model whose belief is
being updated.
