\section{Appendix}

\subsection{Proof Sketch for Theorem 1}

Under a strictly proper scoring rule, the Bayes action is the posterior
predictive distribution. Once the latent reliability variable $r$ is introduced
and the answer posterior is parameterized as a reliability-aware interpolation
between parametric and contextual beliefs, the optimal action is the integral
over posterior reliability:
\[
p^\star(y \mid q,c)=\int p(y \mid q,c,r)\,p(r \mid q,c)\,dr.
\]
Restricting to the implemented log-linear family yields the decoder used by the
Bayes proxy experiments. The missing proof step in the current draft is the
identifiability lemma tying posterior reliability to observable signals.

\subsection{Theorem-1 Log-Linearity Robustness Sweep}

Theorem 1 is exact inside the log-linear family, so we also ran an
11-point synthetic misspecification sweep that perturbs the oracle away from
that family with an explicit nonlinear term while keeping the same
calibration-fit procedure. At calibration size `2048', the fitted rule still
recovers the exact-family oracle nearly perfectly at zero violation (mean
held-out $R^2 = 0.9988$, mean excess Brier $= 0.000154$). More importantly,
degradation is smooth rather than catastrophic under controlled
misspecification: at measured violation strengths `0.3', `0.5', and `0.7' the
held-out $R^2$ remains `0.9891', `0.9609', and `0.8858', respectively, and
even the maximal sweep point only falls to `0.7290'. The
intended empirical interpretation is therefore not that real LLM logits must be
exactly log-linear, but that the implemented Bayes-style rule tracks the
oracle family very closely when the family is approximately right and degrades
gracefully as violations grow.

\subsection{Proof Sketch for Theorem 2}

Take two conflict families, $D_+$ and $D_-$, such that context is respectively
reliable and misleading. Any fixed policy must fail on at least one of these
families. A Yao-style mixture over $\{D_+,D_-\}$ therefore induces a constant
lower bound on the best fixed policy. An adaptive policy that sees a separating
signal $s(q,c)$ incurs regret proportional to the signal error. This establishes
that fixed trust policies are minimax-suboptimal over the conflict family.

\subsection{Theorem-3 Reporting Convention}

The theorem-3 tables report the confidence--accuracy gap rather than only ECE.
For the conflict slices considered here, the gap is the cleanest expression of
the phenomenon of interest: confidence can rise sharply even when accuracy is
flat or falling. We still report ECE in the generated artifacts, and in some
slices the two are numerically very close.

\subsection{Current 14B Scaling Snapshot}

The finished DeepSeek-R1-Distill-Qwen-14B run contains `4200` rows. Its
headline conflict-slice numbers are:

\begin{itemize}
\item WikiContradict conflict: $0.2717 \rightarrow 0.4516 \rightarrow 0.3750$
\item ConflictBank conflict: $0.5876 \rightarrow 0.9449 \rightarrow 0.9513$
\end{itemize}

The first row reinforces the intermediate-CoT peak. The second shows that some
conflict regimes remain catastrophically overconfident even at long CoT. This
is why the final theorem-3 statement is family-dependent rather than uniformly
non-monotone.

\subsection{Answer-Stability Diagnostic for Theorem 3(a)}

The monotone-gap implication in theorem 3(a) is conditional on answer-state
stability across the local CoT window. We therefore measured how often the
predicted answer stays fixed across the canonical `cot=0', `cot=128', and
`cot=1024' checkpoints on the finished
DeepSeek-R1-Distill-Qwen-14B real-generation run.

On `ConflictBank' conflict, answer stability is rare but non-zero:
\begin{itemize}
\item stable across `0/128/1024': $0.08$
\item stable across the pre-saturation window `0/128': $0.098$
\item stable across `128/1024': $0.364$
\end{itemize}

This means theorem 3(a) should be read as a conditional statement about the
stable subset rather than as a universal per-example claim. Importantly, the
stable pre-saturation subset still exhibits the predicted confidence-gap growth:
among `ConflictBank' conflict examples whose answer is unchanged from `cot=0'
to `cot=128', the mean gap increase is $+0.0908$ with zero mean accuracy
change. On the fully stable `0 \rightarrow 1024' subset, the mean gap increase
remains $+0.1025$ with zero mean accuracy change.

\subsection{Benchmark-Dependent Separation Remark}

The theorem-3 threshold should be understood as benchmark-dependent through the
effective self-confirming separation parameter $\delta_c(s)$ rather than as a
single model-only constant. Empirically, `WikiContradict' and `ConflictBank'
fall into different regimes at the same model scale: on the finished 14B run,
`WikiContradict' conflict shows partial long-CoT recovery
($0.2717 \rightarrow 0.4516 \rightarrow 0.3750$), while `ConflictBank'
conflict remains saturated
($0.5876 \rightarrow 0.9449 \rightarrow 0.9513$). The cleanest interpretation
is that controlled entity-swap conflicts induce a larger effective
self-confirming separation than naturalistic contradiction, so the same model
can lie on opposite sides of the safe-update boundary on the two benchmarks.

\subsection{Scope Note on Theorem-3 Rigor}

Theorem 3 should be read differently from the first two results. Theorems 1
and 2 are conventional theorem targets with direct decision-theoretic objects:
posterior-predictive Bayes risk in theorem 1 and a fixed-policy minimax lower
bound in theorem 2. Theorem 3 instead combines a structural analogy from
misspecified-learning theory with direct empirical diagnostics on real traces.
In particular, the Berk--Nash convergence step is ported from the econometrics
and misspecified-learning literature as an analogy for endogenous evidence in
autoregressive decoding, not as a literal transformer-specific convergence
proof. Likewise, the empirical proxy $\hat{\eta}_{\mathrm{eff}}$ is intended as
a directional indicator of safe-versus-unsafe update strength rather than as a
point-identifying estimator of the underlying generalized-Bayes learning rate.

This distinction matters for the cross-family interpretation as well. The
DeepSeek-R1-Distill family gives the clearest saturated-conflict read, but the
same 7B-to-14B asymmetry does not replicate cleanly in Qwen through 32B. That
is why the paper's final theorem-3 statement is benchmark-dependent and
training-family-conditioned rather than a universal scale law. It is also why
we avoid overselling the result as a clean GRPO-versus-DPO causal isolation:
R1-Distill is a distillation product rather than a from-scratch GRPO checkpoint
matched to a DPO baseline.

Three additional scope limits follow from this reading. First, the current
paper does not yet contain a completed head-to-head number against
RLCR-style reward redesign \citep{damani2025rlcr}; the present training-time
evidence is lighter-weight. Second, the retrieval deployment demo is bounded
rather than production-scale. Third, the current theorem-3 evidence is
English-only.

\subsection{Eta-Tempered Decoding Recipe}

The concrete intervention implied by theorem 3 is confidence-only
$\eta$-tempering of the post-trace answer distribution. The recommended recipe
is:
\begin{enumerate}
\item hold out a small calibration split for each model-and-benchmark cell;
\item generate standard CoT traces once;
\item sweep $\eta \in [0.1, 1.0]$ over the post-trace answer distribution;
\item choose the largest $\eta$ that improves Brier score without harming
accuracy on the calibration split;
\item treat large residual gaps after tempering as answer-level failures rather
than calibration-only failures.
\end{enumerate}

The finished 14B intervention read gives the operational headline directly:
conflict slices tolerate only about $0.52\times$ the do-no-harm $\eta$ of
no-conflict slices. WikiContradict conflict is almost fully recalibratable by
confidence-only tempering, while ConflictBank conflict is not.

\subsection{Benchmark-Family Consistency Read}

The overall spotlight-matrix bootstrap interval is the main uncertainty
statement in the paper, but the per-benchmark family breakdown is also
informative. On the expanded spotlight matrix, ConflictBank, FaithEval,
MemoTrap, and NQ-Swap are unanimous $5/5$ Bayes-over-heuristic wins across
model families, while PopQA is the lone mixed family because of the Llama-70B
outlier slice. On the theorem-3 proxy matrix, AmbigDocs, ConflictBank,
FaithEval, and RAMDocs are unanimous $5/5$ Bayes-over-heuristic wins, while
WikiContradict is a unanimous negative exception. This is the cleanest compact
summary of why theorem 3 should be written as benchmark-dependent rather than
universally positive at the regret layer.

\subsection{Named-Baseline Holm Table}

Appendix table `K' extends the benchmark-family sign-test read to every named
baseline in the spotlight matrix: `Self-RAG', `Astute-RAG', `AdaCAD',
`MADAM-RAG', `JuICE', `NWCAD', `CRAG', `CAD', and `CoCoA'. The strictest read
remains the same one highlighted in the body. For `CoCoA', the four benchmark
families `ConflictBank', `FaithEval', `MemoTrap', and `NQ-Swap' are all
unanimous $5/5$ Bayes-over-CoCoA wins with nominal one-sided sign
$p = 0.0312$, which becomes Holm-adjusted $p = 0.1560$ across the five
benchmark families; `PopQA' remains the lone mixed exception. The full table is
generated directly from the completed spotlight report so the multiple-testing
correction is explicit rather than implicit.
