\documentclass{article}

% --- pass options to natbib BEFORE neurips style loads it ---
\PassOptionsToPackage{numbers,sort&compress}{natbib}

\IfFileExists{neurips_2026.sty}{\usepackage[final]{neurips_2026}}{\usepackage[margin=1in]{geometry}}

\usepackage{microtype}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{mathtools}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{array}
\usepackage[dvipsnames]{xcolor}
\usepackage{colortbl}
\usepackage{bm}
\usepackage{tabularx}
\usepackage{enumitem}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{placeins}
\usepackage{afterpage}
\usepackage[colorlinks=true,linkcolor=NavyBlue,citecolor=NavyBlue,urlcolor=NavyBlue]{hyperref}
\usepackage[capitalize,noabbrev]{cleveref}
\usepackage{tcolorbox}
\tcbuselibrary{skins,breakable}

% --- thin frame for emphasized (core) theorem statements ---
\newtcolorbox{ctbox}{enhanced,breakable,colback=white,colframe=black,boxrule=0.4pt,arc=0pt,outer arc=0pt,left=3pt,right=3pt,top=0pt,bottom=0pt,boxsep=1pt,before skip=2pt,after skip=2pt}

% --- theorem environments (separate counter for theorems vs the rest) ---
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}
\newtheorem{example}{Example}

% --- macros ---
\newcommand{\KL}{\operatorname{KL}}
\newcommand{\JSD}{\operatorname{JSD}}
\newcommand{\TV}{\operatorname{TV}}
\newcommand{\softmax}{\operatorname{softmax}}
\newcommand{\one}[1]{\mathbb{1}\!\left[#1\right]}
\newcommand{\E}{\mathbb{E}}
\newcommand{\Var}{\operatorname{Var}}
\newcommand{\PP}{\mathbb{P}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\Y}{\mathcal{Y}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\A}{\mathcal{A}}
\newcommand{\Hh}{\mathcal{H}}
\newcommand{\T}{\mathcal{T}}
\newcommand{\pp}{\mathbf{p}}
\newcommand{\bayes}{\textsc{BayesArbiter}}
\newcommand{\param}{\mathrm{par}}
\newcommand{\ctx}{\mathrm{ctx}}
\newcommand{\closed}{\mathrm{cb}}
\newcommand{\reli}{r}
\newcommand{\good}{\textcolor{ForestGreen}{\textbf{$\bm{\checkmark}$}}}

\newcommand{\headrowbg}{\rowcolor[gray]{0.92}}
\newcommand{\altrow}{\rowcolor[gray]{0.97}}

\floatname{algorithm}{Algorithm}
\algrenewcommand\algorithmicrequire{\textbf{Input:}}
\algrenewcommand\algorithmicensure{\textbf{Output:}}
\makeatletter
\renewcommand{\theHsection}{\thesection}
\renewcommand{\theHsubsection}{\thesubsection}
\renewcommand{\theHequation}{\theequation}
\renewcommand{\theHfigure}{\thefigure}
\renewcommand{\theHtable}{\thetable}
\renewcommand{\theHtheorem}{\thetheorem}
\renewcommand{\theHlemma}{\thelemma}
\renewcommand{\theHproposition}{\theproposition}
\renewcommand{\theHcorollary}{\thecorollary}
\renewcommand{\theHdefinition}{\thedefinition}
\renewcommand{\theHassumption}{\theassumption}
\renewcommand{\theHremark}{\theremark}
\renewcommand{\theHexample}{\theexample}
\renewcommand{\theHALG@line}{\thealgorithm.\arabic{ALG@line}}
\makeatother
\setlength{\textfloatsep}{10pt plus 2pt minus 2pt}
\setlength{\floatsep}{8pt plus 2pt minus 2pt}
\setlength{\intextsep}{8pt plus 2pt minus 2pt}
\captionsetup{font=small,labelfont=bf}
\captionsetup[algorithm]{font=small,labelfont=bf}
\raggedbottom

\title{When Should LLMs Trust Retrieval?\\
       Bayes-Optimal Knowledge Arbitration under Conflict}

\author{%
  Anonymous Authors%
}

\begin{document}
\maketitle

\begin{abstract}
A retrieval-augmented LLM has to arbitrate three voices that often disagree: a parametric prior fixed at training, a retrieved passage, and the model's own chain of thought. Production systems use hand-tuned rules for this; the choice silently fixes hallucination rate and calibration on every slice where the sources disagree, and no principled account has existed. We provide one in three theorems and two inference-time interventions that fall out of a single free-energy functional. \textbf{(CT1)} Inside the log-linear family the Bayes-optimal combiner is the reliability-weighted geometric mixture $\hat p_{w^\star}\!\propto\!p_\theta^{1-w^\star}\,p_{\ctx}^{w^\star}$ with $w^\star\!=\!\E[r\mid q,c]$. It is the unique Gibbs minimizer of an arbitration--tempering free energy and the unique combiner under four classical axioms; CAD, AdaCAD, CoCoA, Astute, Self-RAG, and JuICE all collapse to special cases. The same statement carries a TV-perturbation envelope on regret. \textbf{(CT2)} The chain of thought reads as an endogenous generalized-Bayes update on the same functional. Its rate of convergence to the Berk--Nash fixed point lives inside an $\eta$-linear sandwich, and the safe learning rate shrinks under conflict on the stable-answer subset. \textbf{(CT3)} When the post-trace distribution overshoots, the Brier improvement under $\eta$-tempering is closed-form, $\Delta\mathrm{Brier}(\eta^\star)\!=\!(1\!-\!\eta^\star)^{2}V_{\mathrm{post}}\!+\!O((1\!-\!\eta^\star)^{3})$, giving safe-Bayes an operational reading of when ordinary temperature scaling is justified. The Bayes rule beats the strongest heuristic by $+0.083$ in regret on $23/25$ cells (e-value $2806$); powered $n\!=\!210$ head-to-heads against CoCoA, Astute~RAG, and Self-RAG land at $+0.043 / +0.032 / +0.022$ regret, every CI strictly above zero. The CT2 spectral diagnostic predicts trace decay $\TV(p_t,p_K)$ within $\pm 0.07$ across $12$ cells over $3$ model families (pooled $r\!=\!0.997$). On the worst standalone cell (R1-14B / ConflictBank / $K\!=\!1024$), $\eta$-tempering drops Brier from $0.903$ to $0.505$ and lifts accuracy from $0.037$ to $0.440$. A matched-base contrast (DPO vs.\ GRPO at $8$B, $+0.344$ ECE gap) gives causal evidence the failure mode is RL-induced. On the parallel RLCR-trained $8$B cell, stacking $\eta$-tempering on top reaches accuracy $0.531$ and ECE $0.431$. Both interventions ship as one inference-time line over a standard RAG decoder.
\end{abstract}

\input{sections/introduction}
\input{sections/related_work}
\input{sections/method}
\input{sections/experiments}
\input{sections/discussion}

\input{sections/neurips_checklist}

\clearpage
{\small
\bibliographystyle{plainnat}
\bibliography{references}
}

\appendix
\clearpage
\onecolumn
\input{appendix}

\end{document}
